---
title: "Homework 2"
author: Fiona Ehrich
output: github_document
---

```{r libraries}
library(tidyverse)
library(readxl)
```

## Problem 1

Read the Mr. Trash Wheel dataset.

```{r}
trashwheel_df =
  read_xlsx(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "Mr. Trash Wheel",
    range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
```

Read precipitation data for 2018 and 2017.

```{r}
precip_2018 =
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate(year)

precip_2017 =
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2017 Precipitation",
    skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate(year)
```

Now combine annual precipitation dataframes. The technique being used (creating a "helper" tibble) was presented by Jeff during the live session.

```{r}

month_df =
  tibble(
    month = 1:12,
    month_name = month.name
  )

precip_df =
  bind_rows(precip_2018, precip_2017)

precip_df =
  left_join(precip_df, month_df, by = "month") %>% 
  select(year, month_name, total, -month) # Reordering the variables to be in a more useful order and getting rid of the old month variable
```

**Mr. Trash Wheel data** (`trashwheel_df`)

This dataset contains information from the Mr. Trash Wheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trash wheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, including some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` rows in our final dataset.

The median number of sports balls found in a dumpster in 2017 was `r trashwheel_df %>% filter(year == 2017) %>% pull(sports_balls) %>% median()`.

**Preciptation data** (`precip_df`)

This dataset contains the total preceiptation per month for 2017 and 2018. There are a total of `r nrow(precip_df)` rows in our final dataset.

The total precipitation in 2018 was `r precip_df %>% filter(year == 2018) %>% pull(total) %>% sum()` inches.

## Problem 2

Read and clean the NYC Transit data (including converting the entry variable from character to logical).

```{r}
transit_df =
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>%
  select( # Selecting the desired variables
    line,
    station_name,
    station_latitude,
    station_longitude,
    route1:route11,
    entry,
    vending,
    entrance_type,
    ada
    ) %>% 
  mutate(
    entry = ifelse(entry == "YES", TRUE, FALSE), # Converting the entry variable from character to logical
    vending = ifelse(entry == "YES", TRUE, FALSE), # Converting the vending variable from character to logical
    route8 = as.character(route8), # Converting the route variables that are numeric to character
    route9 = as.character(route9),
    route10 = as.character(route10),
    route11 = as.character(route11)
    )
```

This dataset provides various types of information regarding NYC subway stations and their entrances/exits. The variables in the dataset are: line, station name, station latitude, station longitude, its routes 1 through 11 (each route captured as its own variable), whether or not there is entry, entrance type, and whether or not it is ADA compliant. So far, I have read the dataset, reformatted the names using `janitor::clean_names()`, selected the variables that I would like to keep, converted the `entry` and `vending` variables from character to logical, and converted variables `route 8` through `route 11` from numeric to character. There are `r nrow(transit_df)` rows and `r ncol(transit_df)` columns in the dataset. These data could be tidier. In particular, the route information is displayed in a "wide" fashion and could be combined into two variables (one that expresses route number and one that expresses route name).

There are `r transit_df %>% select(station_name, route1:route11) %>% distinct() %>% nrow()` distinct stations.

`r transit_df %>% filter(ada == TRUE) %>% select(station_name, route1:route11) %>% distinct() %>% nrow()` stations are ADA compliant.

`r (transit_df %>% filter(vending == FALSE & entry == TRUE) %>% nrow())/(transit_df %>% filter(vending == FALSE) %>% nrow())*100`% of the station entrances/exits without vending allow entrance.

